TacticToe: Learning the Game of Tactics

### 2018-01-16 tactictoe_jv.tex
Adding TODOs to the paper

\begin{abstract}
Formalization of mathematical statements in interactive theorem provers (ITPs) 
relies nowadays on many methods to automate the construction of a proof.
Some of this method are general purpose and with enough time can prove any 
theorems. Yet, in a formalized proof script 
a mathematician take advantages of a combination of specialized rules, 
theory-based strategy and general ones. All of them are implemented as tactics 
in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
strategies. Sometimes, a strategy can be specialized for only one theory as in 
SMT solving. But it is difficult for an automated theorem prover to rely on 
rules tailored for a particular case. 
Indeed, this would require implementing many specific rules and more 
importantly figuring out in which cases they can be applied.

@@

\section{Introduction}
%\begin{example} Proof automatically generated by \tactictoe for the user given 
%goal
%\small
%\begin{center}
%
%
%\begin{BVerbatim}[commandchars=\\\{\}]
%Goal: ``\(\forall\)l. FOLDL (\(\lambda\)xs x. SNOC x xs) [] l = l``
%Proof:
%  SNOC_INDUCT_TAC THENL
%  [ REWRITE_TAC [APPEND_NIL, FOLDL],
%    ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC]
%      THEN CONV_TAC (DEPTH_CONV BETA_CONV)
%      THEN ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC] ]
%\end{BVerbatim}
%
%
%\end{center}
%
%\end{example}
%
\vspace{100mm}

@@

\vspace{100mm}
\todo{These 2 paragraphs are copied and slightly edited}
Many of the state-of-the-art interactive theorem provers (ITPs) such as 
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.
%Thanks to a very different proof-search procedure, \tactictoe also
%proves many theorems whose proofs are not found by the hammering
%approach.

\subsection{Contributions}
This project is an extension of the work described in \cite{tgckju-lpar17}. We 
explain here the additional contributions and their effect across our system:

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. Intermediate files storing the result of each step are not 
necessary anymore. So, our recording is more complete and fully 
automatized.
\item Existing tactic orthogonalization on recorded data is made stronger by 
also including the internal ATP \metis in the orthogonalization 
competition. Because we manage a database of theorems for argument 
prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. (todo: A first successful evaluation 
function.)
\item The internal ATP \metis used during our proof search is complemented by 
asynchronous calls to \eprover.
\item First evaluation of \tactictoe and \eprover success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

\section{Related Work}\todo{reorganise}
There are several essential components of our work that are comparable to 
previous approaches: tactic-level proof recording, tactic 
selection through machine learning techniques and automatic tactic-based proof 
search. Our work is also related to previous approaches that use machine 
learning to select premises for the ATP systems and guide ATP proof search 
internally.

@@

\section{Proof Recording}\label{sec:recording}
Recording proofs in an LCF-style proof assistant can be done at different levels.
In \holfour all existing approaches relied on modifying the kernel. This was used
either to export the primitive inference steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable for our
purpose of learning proving strategies at the intermediate tactic level. We therefore
discuss recording proofs in an LCF-style proof assistant, with the focus on \holfour
in this section.

@@

Rather than to rely on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST). The AST is then transformed into a globalized AST
where every top-level value is made independent from
the context. \todo{some reason why this is needed? nice recording level} We 
split the proofs into a series of tactic units and wrap them 
with a recording function.
By replaying the proofs produced by the globalized AST, the wrapper can
record the globalized tactic string with its input goal as training examples in
a database.
The combined effect of these steps on a \holfour proof will be illustrated on Example~\ref{ex:running}.

@@

\section{Prediction}\label{s:prediction}

@@

\section{Orthogonalization}

@@

\section{Abstraction}\label{sec:synthesis}
One of the great weakness of the previous version of \tactictoe is that
it could not create its own tactics. Indeed, sometimes no previous tactic is 
adapted for a certain goal and creating a new tactic is necessary.
%Experiments in section, show how different degrees of tactic synthesis could 
%improve the success rate of \tactictoe.
We present a way to create tactics with different arguments
by abstracting them and re-instantiating them using a predictor for that kind 
of argument. In the spirit of the orthogonalization method, we will try to 
create tactics that are more general but have the same effect as the 
ungeneralized one.

@@

\section{Proof Search}\label{sec:proofsearch}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not 
solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in 
which direction.
Our search algorithm is an advanced tree search algorithm that relies on 
evaluation and policy functions to guide the search.

(todo: missing link)
A proof is in general not a path but a tree. This means that our search space 
has two branching factors: 
the choice of a tactic, and the number of goals produced by each tactic.
The first one can be thought as a disjunction of tactics. Indeed, 
if there exists a tactic that solves a goal then it is solved. And 
conversely, the second branching factor is a conjunction of goals. All pending 
goals must be proven in order to prove the parent goal. For example, in the 
case of induction tactics the pending goals are the base case and the inductive 
case.

If we suppose that each tactic produces at most one goal, then the standard 
monte carlo tree search algorithm applies. So we will first discuss how our 
algorithm works in this case. And then explain how we extend it to support 
tactics producing 
multiple goals.

@@

%Search Algorithm AlphaGo Zero uses a much simpler variant of the asynchronous 
%policy and
%value MCTS algorithm (APV-MCTS) used in AlphaGo Fan and AlphaGo Lee.
%Each node s in the search tree contains edges (s, a) for all legal actions a ∈ 
%A(s). Each edge
%stores a set of statistics,
%{N (s, a), W (s, a), Q(s, a), P (s, a)},
%where N (s, a) is the visit count, W (s, a) is the total action-value, Q(s, a) 
%is the mean action-value,
%and P (s, a) is the prior probability of selecting that edge. Multiple 
%simulations are executed in
%parallel on separate search threads. The algorithm proceeds by iterating over 
%three phases (a–c in
%Figure 2), and then selects a move to play (d in Figure 2).
%25Select (Figure 2a). The selection phase is almost identical to AlphaGo Fan 
%12 ; we recapitulate
%here for completeness. The first in-tree phase of each simulation begins at 
%the root node of the
%search tree, s 0 , and finishes when the simulation reaches a leaf node s L at 
%time-step L. At each
%of these time-steps, t < L, an action is selected according to the statistics 
%in the search tree,
%
%a t = argmax Q(s t , a) + U (s t , a) , using a variant of the PUCT algorithm 
%24 ,
%a
%pP
%U (s, a) = c puct P (s, a)
%N (s, b)
%1 + N (s, a)
%b
%where c puct is a constant determining the level of exploration; this search 
%control strategy initially
%prefers actions with high prior probability and low visit count, but 
%asympotically prefers actions
%with high action-value.
%Expand and evaluate (Figure 2b). The leaf node s L is added to a queue for 
%neural network
%evaluation, (d i (p), v) = f θ (d i (s L )), where d i is a dihedral 
%reflection or rotation selected uniformly
%at random from i ∈ [1..8].
%Positions in the queue are evaluated by the neural network using a mini-batch 
%size of 8; the
%search thread is locked until evaluation completes. The leaf node is expanded 
%and each edge (s L , a)
%is initialised to {N (s L , a) = 0, W (s L , a) = 0, Q(s L , a) = 0, P (s L , 
%a) = p a }; the value v is then
%backed up.
%Backup (Figure 2c). The edge statistics are updated in a backward pass through 
%each step
%t ≤ L. The visit counts are incremented, N (s t , a t ) = N (s t , a t ) + 1, 
%and the action-value is updated
%to the mean value, W (s t , a t ) = W (s t , a t ) + v, Q(s t , a t ) =
%W (s t ,a t )
%.
%N (s t ,a t )
%We use virtual loss to ensure
%each thread evaluates different nodes 69 .

@@

\subsection{Optimization}

@@

We now concentrate on defining a reasonable evaluation function that also 
influences node selection in MCTS. Essentially, we estimate how 
interesting is the list of goals of a tactic to determine if we should continue 
searching in this direction. Theoretically, an interesting list of goal is one 
that has the shortest proof, and we could measure the likelyhood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives. So to reduce this number, we limit 
ourselves to feature vectors tested during orthogonalization. We declare a list 
of goals l to be positive, noted $l \in Positive$, if it is has been 
produced by the winner of an orthogonalization competition. We extend the 
definition of a $distance$ to a list of goals by considering the union of the 
features of the goals. The set of the $r$ set of goals closest to a set of 
goal $l$ is noted $Neighbour(r,l)$. We can define the set of positives and 
negatives within a radius $r$ of a 
list of goals $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
Positive \rbrace$.

@@

\section{ATP Integration}

General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
\holyhammer for \holfour~\cite{tgck-cpp15}. The principle is that we will 
create a \holyhammer tactic that can be used during the proof search. Since 
external calls to \holyhammer are computationally expensive (5 seconds at a 
minimum), we will also 
experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm from previously proven theorems $\mathfrak{T}$ and
a short call (0.1 seconds) to the internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

@@

To integrate a general tactic that requires a higher time out such as \eprover, 
we make use of asynchronous calls to prevent the ATP 
from slowing down the search.  The number of parallel threads can be capped.
And if tactictoe or an asynchronous thread proves a goal, all descendants of 
these goal are deactivated terminating associated threads. This frees space for 
more pending threads associated to newly created nodes to start.

@@

\section{Experimental Evaluation}\label{s:experiments}

Disclaimer:
Comparing results from different tables is possible but not advisable since 
conditions of the experiments may vary slightly because of various factors such 
as overhead cost of additional features or the fact that they were run on 
different machines on a local machine or on the server.

\subsection{Methodology} \todo{is copied}
The evaluation imitates the construction of the library: For each theorem only 
the previous
human proofs are known. These are used as the learning base for the predictions.
To achieve this scenario we re-prove all theorems during a modified build of 
\holfour.
As theorems are proved, they are recorded and included in the training examples.
For each theorem we first attempt to run the \tactictoe search with a time 
limit of 5 seconds,
before processing the original proof script.
In this way, the fairness of the
experiments is guaranteed by construction. 
Only previously declared \sml 
variables (essentially tactics, theorems and simpsets) are accessible. 
And for each theorem to be re-proven \tactictoe is only trained on previous 
proofs.

@@

\subsubsection{Feature parameters}
The first experiment concerns the choice of the right kind of features and 
feature scoring 
mechanism. The results are presented in Table~\ref{tab:feature_param}.
We observe that the higher-order features and the feature of the top logical 
structure increase minimally the number of problems solved. The attempted 
length penalty on the total number of features is actually slightly harmful.
We did not turn on the top feature flag in the next experiments because they 
are slightly more expensive too compute.

@@

@@

\section{Proof Presentation}\label{sec:proofdisplay}

@@

\section{Case study}
%Investigating further the different qualities of \tactictoe, we study its 
%generated proof scripts on an example in \texttt{list} theory (see 
%Example~\ref{ex:casestudy}). 
%The theorem to be proven states the equivalence 
%between the fact that a number $n$ is greater than the length of a list 
%$\mathit{ls}$ with the 
%fact that dropping $n$ elements from this list returns an empty list.
%
%The human proof proceeds by induction on $n$ followed by solving both goals 
%using rewrite steps combined with an arithmetic decision procedure. 
%Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
%induction but solve 
%the base case and the inductive case in a different way.
%The base case only needs rewriting using the global simpset in the 
%\tactictoe(NH)
%proof, which is simulated by a call to \metis in the (SH) proof.
%The inductive case should require an arithmetic decision procedure as hinted 
%by 
%the human 
%proof. This is achieved by rewriting using an arithmetic simpset in the second 
%proof. In the first proof however, a rewriting step and case splitting step 
%were used
%to arrive at a point where \metis calls succeed.
%The tactic proof produced by \tactictoe(NH) often looks better than the one 
%discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
%large numbers of premises.

@@

\section{Conclusion}\label{sec:concl}

Future work:
  Promising techniques that could not be evaluated. Orthogonalization of 
  theorems~\cite{ckju-jsc15}. Term predictions.

Cite in future works.
%A first successful use of orthogonalization techniques for theorems appeared in
%a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined 
%from 
%proofs recorded at the kernel level. Because the number of recorded lemmas was 
%in millions, page ranking techniques was used as a filter to extract 
%``useful'' 
%lemmas.

>>>>

\begin{abstract}\todo{CK review + choose title}
Formalization of mathematical statements in interactive theorem provers (ITPs) 
relies nowadays on many methods to automate the construction of a proof.
Some of this method are general purpose and with enough time can prove any 
theorems. Yet, in a formalized proof script 
a mathematician take advantages of a combination of specialized rules, 
theory-based strategy and general ones. All of them are implemented as tactics 
in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
strategies. Sometimes, a strategy can be specialized for only one theory as in 
SMT solving. But it is difficult for an automated theorem prover to rely on 
rules tailored for a particular case. 
Indeed, this would require implementing many specific rules and more 
importantly figuring out in which cases they can be applied.

@@

\section{Introduction}\todo{TG + CK review}
%\begin{example} Proof automatically generated by \tactictoe for the user given 
%goal
%\small
%\begin{center}
%
%
%\begin{BVerbatim}[commandchars=\\\{\}]
%Goal: ``\(\forall\)l. FOLDL (\(\lambda\)xs x. SNOC x xs) [] l = l``
%Proof:
%  SNOC_INDUCT_TAC THENL
%  [ REWRITE_TAC [APPEND_NIL, FOLDL],
%    ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC]
%      THEN CONV_TAC (DEPTH_CONV BETA_CONV)
%      THEN ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC] ]
%\end{BVerbatim}
%
%
%\end{center}
%
%\end{example}
%
\vspace{100mm}

@@

\vspace{100mm}
\todo{TG : comparison with hammers and advantages for the user and over hammers}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todo{TG: rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.
%Thanks to a very different proof-search procedure, \tactictoe also
%proves many theorems whose proofs are not found by the hammering
%approach.

\todo{CK review}
\subsection{Contributions}
This project is an extension of the work described in \cite{tgckju-lpar17}. We 
explain here the additional contributions and their effect across our system:

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. So, our recording is more complete and fully automatized.
\item Existing tactic orthogonalization on recorded data is made stronger by 
also including the internal ATP \metis in the orthogonalization 
competition. Because we manage a database of theorems for argument 
prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently of the tactic. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. The policy and evaluation are learned 
through supervised learning.
\item The internal ATP \metis used during our proof search is complemented by 
asynchronous calls to \eprover.
\item First evaluation of \tactictoe and \eprover success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

\section{Related Work}\todo{TG reorganise + other motivation?}
There are several essential components of our work that are comparable to 
previous approaches: tactic-level proof recording, tactic 
selection through machine learning techniques and automatic tactic-based proof 
search. Our work is also related to previous approaches that use machine 
learning to select premises for the ATP systems and guide ATP proof search 
internally.

@@

\section{Proof Recording}\label{sec:recording}
\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
mechanism and replace by a high level abstraction. Alternatively, make the 
current version more accessible.}

Recording proofs in an LCF-style proof assistant can be done at different levels.
In \holfour all existing approaches relied on modifying the kernel. This was used
either to export the primitive inference steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable for our
purpose of learning proving strategies at the intermediate tactic level. We therefore
discuss recording proofs in an LCF-style proof assistant, with the focus on \holfour
in this section.

@@

Rather than to rely on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
human ==> better for the machine learning algorithm (learn the right thing)}

 The AST is then transformed into a globalized AST
where every top-level value is made independent from
the context. \todo{some reason why this is needed? easy transfer between 
theories} We 
split the proofs into a series of tactic units and wrap them 
with a recording function.
By replaying the proofs produced by the globalized AST, the wrapper can
record the globalized tactic string with its input goal as training examples in
a database.
The combined effect of these steps on a \holfour proof will be illustrated on Example~\ref{ex:running}.

@@

\section{Prediction}\label{s:prediction}\todo{TG: complete term theorem and 
list of goals specificities in terms of predictions and justify by their use 
case}

@@

\section{Orthogonalization}\todo{MN + RK: review + fix typos}

@@

\section{Abstraction}\label{sec:synthesis}\todo{MN + RK: review + fix typos}
One of the great weakness of the previous version of \tactictoe is that
it could not create its own tactics. Indeed, sometimes no previous tactic is 
adapted for a certain goal and creating a new tactic is necessary.
%Experiments in section, show how different degrees of tactic synthesis could 
%improve the success rate of \tactictoe.
We present a way to create tactics with different arguments
by abstracting them and re-instantiating them using a predictor for that kind 
of argument. In the spirit of the orthogonalization method, we will try to 
create tactics that are more general but have the same effect as the 
ungeneralized one.

@@

\section{Proof Search}\label{sec:proofsearch}
\todo
{
TG:
describe the prior policy, the evaluation, 
the node selection algorithm,
the node extension (+ case of solved goal) + backpropagation.
add description of end of the algorithm.
}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not 
solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in 
which direction.
Our search algorithm is an advanced tree search algorithm that relies on 
evaluation and policy functions to guide the search.

\todo{TG: add transition}
A proof is in general not a path but a tree. This means that our search space 
has two branching factors: 
the choice of a tactic, and the number of goals produced by each tactic.
The first one can be thought as a disjunction of tactics. Indeed, 
if there exists a tactic that solves a goal then it is solved. And 
conversely, the second branching factor is a conjunction of goals. All pending 
goals must be proven in order to prove the parent goal. For example, in the 
case of induction tactics the pending goals are the base case and the inductive 
case.

\todo{TG: add general overview of the algorithm and describe the plan}
If we suppose that each tactic produces at most one goal, then the standard 
monte carlo tree search algorithm applies. So we will first discuss how our 
algorithm works in this case. And then explain how we extend it to support 
tactics producing 
multiple goals.

@@

@@

\subsection{Node extension}

To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
(optimized from previous paper)

@@

We now concentrate on defining a reasonable evaluation function that also 
influences node selection in MCTS. Essentially, we estimate how 
interesting is the list of goals of a tactic to determine if we should continue 
searching in this direction. Theoretically, an interesting list of goal is one 
that has the shortest proof, and we could measure the likelihood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives\todo{TG: cite Hard Negative mining (cite 
Deep math)}. So to reduce this number, we limit 
ourselves to feature vectors tested during orthogonalization. We declare a list 
of goals l to be positive, noted $l \in Positive$, if it is has been 
produced by the winner of an orthogonalization competition. We extend the 
definition of a $distance$ to a list of goals by considering the union of the 
features of the goals. The set of the $r$ set of goals closest to a set of 
goal $l$ is noted $Neighbour(r,l)$. We can define the set of positives and 
negatives within a radius $r$ of a 
list of goals $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
Positive \rbrace$.

@@

\section{ATP Integration}\todo{JU + CK: rework and review}

General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
\holyhammer for \holfour~\cite{tgck-cpp15}. The principle is that we will 
create a \holyhammer tactic that can be used during the proof search. Since 
external calls to \holyhammer are computationally expensive (5 seconds at a 
minimum), we  
experiment with a small hammer tactic comprised of: a faster premise 
selection algorithm from previously proven theorems $\mathfrak{T}$ and
a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
internal prover \metis~\cite{metis} or a
longer call to external prover \eprover.

@@

To integrate a general tactic that requires a higher time out such as \eprover, 
we make use of asynchronous calls to prevent the ATP 
from slowing down the search.  The number of parallel threads can be capped.
And if tactictoe or an asynchronous thread proves a goal, all descendants of 
these goal are deactivated, terminating associated threads. This frees space 
for more pending threads, associated to newly created nodes, to start.

@@

\section{Experimental Evaluation}\label{s:experiments}\todo{TG: introduction + 
consistent tables}

Disclaimer:
Conditions of the experiments may vary slightly between different tables 
because of various factors such 
as overhead cost of additional features or the fact that they were run on 
different machines on a local machine or on the server.

\subsection{Methodology} \todo{TG: is copied, update references of sections and
new algorithm + mention when orthogonalization is done}

The evaluation imitates the construction of the library: For each theorem only 
the previous human proofs are known. These are used as the learning base for 
the predictions.
To achieve this scenario we re-prove all theorems during a modified build of 
\holfour.
As theorems are proved, their tactical proofs and their statements are recorded 
and included in the 
training examples.
For each theorem we first attempt to run the \tactictoe search with a time 
limit of 5 seconds,
before processing the original proof script.
In this way, the fairness of the
experiments is guaranteed by construction. 
Only previously declared \sml 
variables (essentially tactics, theorems and simpsets) are accessible. 
And for each theorem to be re-proven \tactictoe is only trained on previous 
proofs.

@@

\subsubsection{Prediction parameters}\todo{TG: remove this section or mention
that parameters were optimized in the previous version}
The first experiment concerns the choice of the right kind of features and 
feature scoring 
mechanism. The results are presented in Table~\ref{tab:feature_param}.
We observe that the higher-order features and the feature of the top logical 
structure increase minimally the number of problems solved. The attempted 
length penalty on the total number of features is actually slightly harmful.
We did not turn on the top feature flag in the next experiments because they 
are slightly more expensive too compute.

@@

\subsubsection{MCTS parameters: policy and evaluation}

@@

\section{Proof Presentation}\label{sec:proofdisplay}\todo{MN + RK: add 
motivations and transitions + review}

@@

\section{Case study}\todo{TG: rerun experiment with prettification and 
minimization + add another example. 
  TG + RK: describe the two examples + find examples that did not appear during 
  evaluation
  RK: After implementation and small experiment is done (in about 1 week) 
      Test tactictoe on new goals and manual exploration mode.
      Give feedback.
}
%Investigating further the different qualities of \tactictoe, we study its 
%generated proof scripts on an example in \texttt{list} theory (see 
%Example~\ref{ex:casestudy}). 
%The theorem to be proven states the equivalence 
%between the fact that a number $n$ is greater than the length of a list 
%$\mathit{ls}$ with the 
%fact that dropping $n$ elements from this list returns an empty list.
%
%The human proof proceeds by induction on $n$ followed by solving both goals 
%using rewrite steps combined with an arithmetic decision procedure. 
%Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
%induction but solve 
%the base case and the inductive case in a different way.
%The base case only needs rewriting using the global simpset in the 
%\tactictoe(NH)
%proof, which is simulated by a call to \metis in the (SH) proof.
%The inductive case should require an arithmetic decision procedure as hinted 
%by 
%the human 
%proof. This is achieved by rewriting using an arithmetic simpset in the second 
%proof. In the first proof however, a rewriting step and case splitting step 
%were used
%to arrive at a point where \metis calls succeed.
%The tactic proof produced by \tactictoe(NH) often looks better than the one 
%discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
%large numbers of premises.

@@

\section{Conclusion}\label{sec:concl}\todo{TG + JU review}

Future work:
  Promising techniques that could not be evaluated. 
  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
  + Try harder on term predictions. \cite{latest version of sepia}.

%A first successful use of orthogonalization techniques for theorems appeared in
%a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined 
%from 
%proofs recorded at the kernel level. Because the number of recorded lemmas was 
%in millions, page ranking techniques was used as a filter to extract 
%``useful'' 
%lemmas.
